Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-07-18 ms.tex
still overhauling MCMC section(s)

For each pixel $p$ and band $\nu_i$, the variance of the measured value 
$I_{\nu_i}(p)$ is taken to be:

\begin{equation}
\sigma_{\nu_i}(p) = c^2_{\nu_i}I_{\nu_i}(p)^2 + c^2_{\nu_i}CMB(p)^2 + (\delta O_{\nu_i})^2 + n_{\nu_i}(p)^2
\end{equation}

@@

With $\bar{T}_2$ = 15.7 K and $\sigma_{\bar{T}_2}$ = 1.4 K. The $T_2$ prior 
mean is chosen based on the typical high-latitude $T_2$ value derived from
the correlation slope analysis from $\S$\ref{sec:global}. We find, as 
desired, that this relatively broad $T_2$ prior has little influence on the 
derived temperatures, other than to regularize the rare pixels with one or more
defective intensities which might otherwise yield unreasonable parameter
estimates. In principle, there can also be an informative prior on 
$\tilde{I}_{545}$. However, we have chosen to assume a uniform prior on the SED
normalization and, as a matter of notation, will omit $P(\tilde{I}_{545})$ 
henceforward. In practice we always perform computations using logarithms of 
the relevant probabilities.
%mention something about the proposal distribution at some point?

@@

For each pixel, we initialize the Markov chain with parameters 
$\tilde{I}_{545}$ = $I_{545}$ and $T_2$ consistent with the FDS99 
DIRBE 100$\mu$m/240$\mu$m ratio map $\mathscr{R}$. The initial proposal 
distribution is a two-dimensional normal distribution, with 
$\sigma_{T_2}$=0.25 K, 
$\sigma_{\tilde{I}_{545}}$=max(0.01$\times$$I_{545}$, 0.05 MJy/sr) and
$\sigma_{T_2}\sigma_{\tilde{I}_{545}}$=0. We run 5 iterations of burn-in, each 
consisting of 500 MH steps. After each burn-in iteration, we rescale the 
proposal distribution so as to ultimately attain an acceptance fraction 
$f_{acc}$ as close as possible to the optimal value $f_{opt}=0.234$. This is 
accomplished by multiplying the proposal distribution standard deviations by 
$f_{acc}$/$f_{opt}$.

@@

After obtaining this initial estimate of the covariance matrix for each
pixel, we re-run the entire MCMC procedure, starting from the first burn-in 
iteration, but this time beginning with a proposal distribution that is a 
two-dimensional normal distribution with covariance equal to the first-pass
covariance estimate. This gives the proposal distribution for each pixel
 approximately the ``right shape'', whereas on the first pass we had to start
by making ad hoc assumptions about the relative widths of the proposal 
distribution in $\tilde{I}_{545}$ and $T_2$, and also assumed that the 
$\sigma_{T_2}\sigma_{\tilde{I}_{545}}$=0 for the proposal distribution. We
validated the recovered parameters and uncertainties from our MCMC procedure
by comparing with results based on gridded posterior calculations. These 
comparisons verified that the proposal distribution rescaling and reshaping 
steps that we employ do improve the accuracy of the recovered 
parameters/uncertainties, and that the parameters/uncertainties ultimately 
derived are highly reliable.

Lastly, during post burn-in sampling, we also estimate the monochromatic 
two-component intensity at 545 GHz, $M_{545}$ = $\langle M_{545, j} \rangle$ = 
$\langle \tilde{I}_{545, j}/b_{545}(T_{2,j}) \rangle$, its variance, and
the 545 GHz optical depth $\langle \tau_{545, j} \rangle$ = 
$\langle M_{545, j}/S_{545, j} \rangle$ and its variance. $\tau_{545}$ and 
$M_{545}$ are more readily useful than the sampling parameters themselves for 
translating our fit results into predicted reddening ($\S$\ref{sec:ebv}) and 
thermal dust emission ($\S$\ref{sec:lofreq}), respectively.

%$M_{545}$ is of more general utility in making foreground predictions because 
%it provides the normalization for the two-component spectrum $M_{\nu}$, 
%independent of the \PLANCK~545 GHz bandpass. Finally, we acompute the 545 GHz
%optical depth as $\tau_{545}$ = and its variance, with the source 
%function $S_{545}$ calculated according to Equation \ref{eqn:source}.

\subsection{Low-resolution Fits}
\label{sec:lores}
%mention that the pixelization changes ---> to lower HEALpix nside
%mention that over some region of the sky with high signal and high
%S/N we try fitting t2, normalization and one of q1/q2, f_1
%using more bands (including 143, and maybe 100 GHz)

%As explained in $\S$\ref{sec:modeling} and $\S$\ref{sec:samp}, our 
%full-resolution fits fix the parameters $f_1$, $q_1/q_2$, $\beta_1$, $\beta_2$ 
%globally so that only the dust temperatures and SED normalization vary on a 
%per-pixel basis. The reason for fixing so many parameters is that, at $6.1'$
%FWHM, only five SED measurements are available per pixel.

% need some sort of add on to last sentence or another sentence entirely
% to the effect of ``staying in overconstrained regime''

>>>>

For each pixel $p$ in band $\nu_i$, the variance of the measured value 
$I_{\nu_i}(p)$ is taken to be:

\begin{multline}
\sigma^2_{\nu_i}(p) = c^2_{\nu_i}I_{\nu_i}(p)^2 + c^2_{\nu_i}\sigma_{CMB}^2 + (\delta O_{\nu_i})^2 \\
+ n_{\nu_i}^2 + \sigma^2_{\nu_i, CIBA} + \sigma^2_{CO}(p)
\end{multline}

@@

With $\bar{T}_2$ = 15.7 K and $\sigma_{\bar{T}_2}$ = 1.4 K. The $T_2$ prior 
mean is chosen based on the typical high-latitude $T_2$ value derived from
the correlation slope analysis of $\S$\ref{sec:global}. We find, as 
desired, that this relatively broad $T_2$ prior has little influence on the 
derived temperatures, other than to regularize the rare pixels with one or more
defective intensities which might otherwise yield unreasonable parameter
estimates. In principle, there can also be an informative prior on 
$\tilde{I}_{545}$. However, we have chosen to assume a uniform prior on the SED
normalization and, as a matter of notation, will omit $P(\tilde{I}_{545})$ 
henceforward. In practice we always perform computations using logarithms of 
the relevant probabilities.

@@

For each pixel, we initialize the Markov chain with parameters 
$\tilde{I}_{545}$ = $I_{545}$ and $T_2$ consistent with the FDS99 
DIRBE 100$\mu$m/240$\mu$m ratio map $\mathscr{R}$. The initial proposal 
distribution is a two-dimensional normal distribution, with 
$\sigma_{T_2}$=0.25 K, 
$\sigma_{\tilde{I}_{545}}$=max(0.01$\times$$I_{545}$, 0.05 MJy/sr) and
$\rho_{T_2, \tilde{I}_{545}}$=0. We run 5 iterations of burn-in, each 
consisting of 500 MH steps. After each burn-in iteration, we rescale the 
proposal distribution so as to ultimately attain an acceptance fraction 
$f_{acc}$ as close as possible to the optimal value $f_{opt}=0.234$. This is 
accomplished by multiplying the proposal distribution standard deviations by 
$f_{acc}$/$f_{opt}$.

@@

After obtaining this initial estimate of the covariance matrix for each pixel, 
we re-run a second iteration of the entire MCMC procedure, starting from the 
first burn-in period. On this iteration, for each pixel, we begin with a 
proposal distribution that is a two-dimensional Gaussian with covariance equal 
to the first-pass covariance estimate. This gives the each pixel's proposal 
distribution approximately the `right shape', whereas on the first pass we 
started by simply guessing the relative widths of the proposal distribution in 
$\tilde{I}_{545}$, $T_2$, and also assumed that the 
$\rho_{T_2, \tilde{I}_{545}}$=0.

We validated the recovered parameters and uncertainties from our MCMC procedure
by comparing with results based on gridded posterior calculations. These 
comparisons verified that the proposal distribution rescaling and reshaping 
steps that we employ do improve the accuracy of the recovered 
parameters/uncertainties, and that the parameters/uncertainties ultimately 
derived are highly reliable.

Lastly, during post burn-in sampling, we also estimate the monochromatic 
two-component intensity at 545 GHz, $M_{545}$ = $\langle M_{545, j} \rangle$ = 
$\langle \tilde{I}_{545, j}/b_{545}(T_{2,j}) \rangle$, its variance, and the 
545 GHz optical depth $\tau_{545}$=$\langle \tau_{545, j} \rangle$ = 
$\langle M_{545, j}/S_{545, j} \rangle$ and its variance. $\tau_{545}$ and 
$M_{545}$ are more readily useful than the sampling parameters themselves for 
translating our fit results into predictions of reddening ($\S$\ref{sec:ebv}) 
and thermal dust emission ($\S$\ref{sec:lofreq}), respectively.

\subsection{Low-resolution Fits}
\label{sec:lores}
